---
title: Data Visualizations
#author:
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document:
    reference_docx: ../../templates/template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      purl = TRUE,
                      out.extra = '',
                      tab.cap.pre = "Table ", tab.cap.sep = ": ",
                      knitr.duplicate.label = "allow",
                      tidy.opts = list(width.cutoff = 40), 
                      tidy = TRUE,
                      strip.white = FALSE)

```

# Install and load libraries

We need to first install all the required packages for the data analysis process. If you have never used the packages below, it is more likely that you have not installed them on your machine either. Please make sure you install each of the packages below using the following command:

```{r eval=FALSE, warning=FALSE, message=FALSE}
install.packages("here")
install.packages("tidyverse")
install.packages("janitor")
install.packages("knitr")
install.packages("flextable")
install.packages("officer")
install.packages("extrafont")
# font_import() # Only do this once
```

Then you need to load the following packages:

```{r load libraries, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(janitor)
library(knitr)
library(flextable)
library(officer)
library(extrafont)
loadfonts(device = "win")
```

# Import clean dataset

```{r import dataset, warning=FALSE, message=FALSE}
unfccc_emissions <- read.csv(here("scripts/cleaning", "unfccc-emissions-clean.csv"))
```

# Data visualizations

Now that we've analyzed the data it's time to produce some visualizations. Given that there are many missing values within the Non-Annex I countries, we should avoid making any inference about their emissions based on this dataset. However, we can visualize Annex I countries' cumulative and annual emissions and figure the most contributing emitters based on their region.

First, let's explore the highest emitting regions in 2018.

```{r}



```


