---
title: UNFCCC Emissions Data Visualization Tutorial
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
output:
  word_document:
    reference_docx: ../../../templates/template.docx
  pdf_document: 
    keep_tex: true
    highlight: pygments
    includes:
      in_header: "latex-header.tex"
bibliography: packages.bib
nocite: '@*'
csl: ../../../bibliography/chicago-fullnote-bibliography-with-ibid.csl
compact-title: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      purl = TRUE,
                      out.extra = '',
                      tab.cap.pre = "Table ", tab.cap.sep = ": ",
                      knitr.duplicate.label = "allow",
                      tidy.opts = list(width.cutoff = 40), 
                      tidy = TRUE,
                      strip.white = FALSE)

```

# Install and load libraries

We need to first install all the required packages for the data analysis process. If you have never used the packages below, it is more likely that you have not installed them on your machine either. Please make sure you install each of the packages below using the following command:

```{r eval=FALSE, warning=FALSE, message=FALSE}
install.packages("here")
install.packages("janitor")
#install.packages("tidyverse")
install.packages("dplyr")
install.packages("stringr")
install.packages("ggplot2")
install.packages("treemapify")
install.packages("knitr")
install.packages("extrafont")
# font_import() # Only do this once
```

Then you need to load the following packages:

```{r load libraries, warning=FALSE, message=FALSE}
library(here)
library(janitor)
#library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(scales)
library(treemapify)
library(knitr)
library(extrafont)
loadfonts(device = "win")
```

# Import clean dataset

```{r import dataset, warning=FALSE, message=FALSE}
unfccc_emissions <- utils::read.csv(here("scripts/cleaning/unfccc-emissions", "unfccc-emissions-clean.csv"), stringsAsFactors = TRUE)
```

# Data visualizations

Now that we've analyzed the data it's time to produce some visualizations. Given that there are many missing values within the Non-Annex I countries, we should avoid making any inference about their emissions based on this dataset. However, we can visualize Annex I countries' cumulative and annual emissions and figure the most contributing emitters based on their region.

First, let's explore the highest emitting regions in 2018.


```{r ghg 2018 by region, warning=FALSE, message=TRUE, caption = "Annex I GHG emissions without LULUCF (2018)"}
unfccc_emissions %>%
  dplyr::filter(year == "2018", group == "Annex I", !(country %in% c("European Union (Convention)", "European Union (KP)"))) %>%
  dplyr::group_by(region, type) %>%
  dplyr::summarise(total = sum(ghg)) %>%
  ggplot2::ggplot(aes(x = as.factor(region), y = total, fill = stringr::str_wrap(as.factor(type), 12))) +
  ggplot2::scale_fill_brewer(palette = "Paired", name = "Type") +
  ggplot2::geom_bar(stat="identity", position=position_dodge(), width = 1) +
  ggplot2::geom_text(aes(label = round(total/1000000, digits = 2)), vjust = -0.2, position=position_dodge(width=0.9)) +
  ggplot2::theme_minimal() +
  ggplot2::xlab("Regions") +
  ggplot2::ylab("GHG (Gt)") +
  ggplot2::scale_y_continuous(labels = scales::unit_format(unit = "", scale = 1e-6)) +
  ggplot2::scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
#  theme(legend.position = "none") +
  ggplot2::theme(text=element_text(family = "Arial", size = 11), 
                 axis.title.x = element_text(family = "Arial", size = 12), 
                 axis.title.y = element_text(family = "Arial", size = 12), 
                 legend.text=element_text(family = "Arial", size=10))
```


When we graph by region and whether it's produced or net emissions (with/without LULUCF), we see that net emissions give Annex I countries the chance to report lower emissions. We can also see that the majority of emissions are located in Europe and Central Asia and North America for Annex I countries so we need to interrogate why that is.

```{r ghg 2018 by region treemap, warning=FALSE, message=TRUE, caption = "Annex I GHG emissions without LULUCF (2018)"}
unfccc_emissions %>%
  dplyr::filter(year == "2018", group == "Annex I", type == "Total GHG emissions without LULUCF", !(country %in% c("European Union (Convention)", "European Union (KP)"))) %>%
  dplyr::mutate(percent = round(ghg/sum(ghg) * 100, digits = 2)) %>%
  dplyr::group_by(region, type) %>%
  ggplot2::ggplot(aes(area = ghg, fill = region, subgroup = region, label = paste(country, paste(round(ghg/1000000, digits = 2), " Gt (", percent, "%)", sep = ""), sep = "\n"))) +
  ggplot2::scale_fill_brewer(palette = "Paired") +
  treemapify::geom_treemap() +
  treemapify::geom_treemap_subgroup_border(size = 0) +
  treemapify::geom_treemap_subgroup_text(grow = FALSE, place = "bottomleft", size = 14, color = "grey20", reflow = TRUE, family = "Arial") +
  ggplot2::theme_void() + 
  treemapify::geom_treemap_text(colour = "white", place = "topleft", reflow = T, size = 8, family = "Arial") +
  ggplot2::theme(legend.position = "none")

ggplot2::ggsave(here("images", "annex-i-ghg-2018.svg"), device="svg", dpi=300)
```



Now let's see if it's the same story when we look at the cumulative GHG emissions from 1990 to 2018.


```{r ghg cumulative by region treemap, warning=FALSE, message=TRUE, caption = "Annex I cumulative GHG emissions without LULUCF (1990-2018)"}
unfccc_emissions %>%
  dplyr::filter(year != "base_year", group == "Annex I", type == "Total GHG emissions without LULUCF", !(country %in% c("European Union (Convention)", "European Union (KP)"))) %>%
  dplyr::group_by(iso, country, region) %>%
  dplyr::summarise_if(is.numeric, sum, na.rm = TRUE) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(percent = round(ghg/sum(ghg) * 100, digits = 2)) %>%
  dplyr::group_by(country, region) %>%
  ggplot2::ggplot(aes(area = ghg, fill = region, subgroup = region, label = paste(country, paste(round(ghg/1000000, digits = 2), " Gt (", percent, "%)", sep = ""), sep = "\n"))) +
  ggplot2::scale_fill_brewer(palette = "Paired") +
  treemapify::geom_treemap() +
  treemapify::geom_treemap_subgroup_border(size = 0) +
  treemapify::geom_treemap_subgroup_text(grow = FALSE, place = "bottomleft", size = 14, color = "grey20", reflow = TRUE, family = "Arial") +
  ggplot2::theme_void() + 
  treemapify::geom_treemap_text(colour = "white", place = "topleft", reflow = T, size = 8, family = "Arial") +
  ggplot2::theme(legend.position = "none")

ggplot2::ggsave(here("images", "annex-i-ghg-cumulative.svg"), device="svg", dpi=300)
rm(unfccc_emissions)
```


# Export as an R script for future use

Only run this chunk manually once within the .Rmd file. It produces an error when knitting it as a whole because of chunk label duplicates. As of `r format(Sys.time(), '%B %d, %Y')`, there hasn't been a viable solution to run the code below when as part of the knitting process.

```{r export as an R script, eval=FALSE, message=FALSE, warning=FALSE}
knitr::purl("unfccc-emissions-visualizations.Rmd", "unfccc-emissions-visualizations.R")
knitr::write_bib(.packages(), "packages.bib")
```

# Software used
