---
title: Data Importing and Wrangling
#author:
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document:
    reference_docx: ../../Templates/template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      purl = TRUE,
                      out.extra = '',
                      knitr.duplicate.label = "allow",
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)

```

# Install and load libraries

```{r load libraries, warning=FALSE, message=FALSE}
library(openxlsx)
library(tidyverse)
library(janitor)
library(knitr)
```

- `openxlsx`: 

- `tidyverse`:

- `janitor`:

- `knitr`:

# Data Import and Wrangling


## Import data


### Annex I parties

```{r import data (Annex I)}
Annex_I_emissions <- openxlsx::read.xlsx("Annual-Net-emissions-removals-Annex-I-raw.xlsx",
                                         startRow = 5, 
                                         fillMergedCells = TRUE, 
                                         rows = 5:96) %>% 
  janitor::clean_names()
```


### Non-Annex I parties

```{r import data (Non-Annex I)}
Non_Annex_I_emissions <- openxlsx::read.xlsx("Annual-Net-emissions-removals-Non-Annex-I-raw.xlsx",
                                             startRow = 5, 
                                             fillMergedCells = TRUE, 
                                             rows = 5:96) %>% 
  janitor::clean_names()
```

## Clean variables

### Annex I parties

```{r clean main dataframe (Annex I), warning=FALSE, message=FALSE}
Annex_I_emissions <- Annex_I_emissions[-c(1),]

colnames(Annex_I_emissions)[1] <- "party"

Annex_I_emissions <- Annex_I_emissions %>% 
  dplyr::rename_all(funs(str_replace_all(., "gh_gs", "ghg")))
```

### Non-Annex I parties

```{r clean main dataframe (Non-Annex I), warning=FALSE, message=FALSE}
Non_Annex_I_emissions <- Non_Annex_I_emissions[-c(1),]

colnames(Non_Annex_I_emissions)[1] <- "party"

Non_Annex_I_emissions <- Non_Annex_I_emissions %>% 
  dplyr::rename_all(funs(str_replace_all(., "gh_gs", "ghg")))
```

## Disaggreagate data by gas type


### Annex I parties

```{r disaggregate dataframe by gas type (Annex I)}
Annex_I_emissions_ch4 <- Annex_I_emissions %>% 
  dplyr::select("party", "gas", contains("ch4")) %>% 
  dplyr::rename("base_year" = "ch4")

colnames(Annex_I_emissions_ch4)[4:32] <- 1990:2018

Annex_I_emissions_co2 <- Annex_I_emissions %>% 
  dplyr::select("party", "gas", contains("co2")) %>% 
  dplyr::rename("base_year" = "co2")

colnames(Annex_I_emissions_co2)[4:32] <- 1990:2018

Annex_I_emissions_ghg <- Annex_I_emissions %>% 
  dplyr::select("party", "gas", contains("ghg")) %>% 
  dplyr::rename("base_year" = "aggregate_ghg")

colnames(Annex_I_emissions_ghg)[4:32] <- 1990:2018
```


### Non-Annex I parties

```{r disaggregate dataframe by gas type (Non-Annex I)}
Non_Annex_I_emissions_ch4 <- Non_Annex_I_emissions %>% 
  dplyr::select("party", "gas", contains("ch4"))

colnames(Non_Annex_I_emissions_ch4)[3:31] <- 1990:2018

Non_Annex_I_emissions_co2 <- Non_Annex_I_emissions %>% 
  dplyr::select("party", "gas", contains("co2"))

colnames(Non_Annex_I_emissions_co2)[3:31] <- 1990:2018

Non_Annex_I_emissions_ghg <- Non_Annex_I_emissions %>% 
  dplyr::select("party", "gas", contains("ghg"))

colnames(Non_Annex_I_emissions_ghg)[3:31] <- 1990:2018
```

## Convert datasets to long

### Annex I parties

```{r convert datasets to long (Annex I)}
Annex_I_emissions_ch4 <- Annex_I_emissions_ch4 %>% 
  tidyr::gather(year, ch4, base_year:`2018`)

Annex_I_emissions_co2 <- Annex_I_emissions_co2 %>% 
  tidyr::gather(year, co2, base_year:`2018`)

Annex_I_emissions_ghg <- Annex_I_emissions_ghg %>% 
  tidyr::gather(year, ghg, base_year:`2018`)
```

### Non-Annex I parties

```{r convert datasets to long (Non-Annex I)}
Non_Annex_I_emissions_ch4 <- Non_Annex_I_emissions_ch4 %>% 
  tidyr::gather(year, ch4, `1990`:`2018`)

Non_Annex_I_emissions_co2 <- Non_Annex_I_emissions_co2 %>% 
  tidyr::gather(year, co2, `1990`:`2018`)

Non_Annex_I_emissions_ghg <- Non_Annex_I_emissions_ghg %>% 
  tidyr::gather(year, ghg, `1990`:`2018`)
```


## Merge datasets back

### Annex I parties

```{r merge datasets back (Annex I)}
Annex_I_emissions <- Annex_I_emissions_ghg %>% 
  dplyr::full_join(Annex_I_emissions_co2, by =c("party", "gas", "year")) %>%
  dplyr::full_join(Annex_I_emissions_ch4, by =c("party", "gas", "year"))

rm(Annex_I_emissions_ch4,Annex_I_emissions_co2,Annex_I_emissions_ghg)
```

### Non-Annex I parties

```{r merge datasets back (Non-Annex I)}
Non_Annex_I_emissions <- Non_Annex_I_emissions_ghg %>% 
  dplyr::full_join(Non_Annex_I_emissions_co2, by =c("party", "gas", "year")) %>% 
  dplyr::full_join(Non_Annex_I_emissions_ch4, by =c("party", "gas", "year"))

rm(Non_Annex_I_emissions_ch4,Non_Annex_I_emissions_co2,Non_Annex_I_emissions_ghg)
```


## Merge Annex I and Non-Annex I

```{r Merge Annex I and Non-Annex I dataframes, message=FALSE, warning=FALSE}
UNFCCC_Emissions <- dplyr::bind_rows(Annex_I_emissions,Non_Annex_I_emissions) %>%
  dplyr::rename("type" = "gas") %>% 
  dplyr::mutate(type = stringr::str_replace(type,"Total GHG emissions excluding LULUCF/LUCF", "Total GHG emissions without LULUCF")) %>% 
  dplyr::mutate(type = stringr::str_replace(type,"Total GHG emissions including LULUCF/LUCF", "Total GHG emissions with LULUCF")) %>%
  readr::type_convert() %>%
  dplyr::mutate_if(is.character, as.factor)

rm(Annex_I_emissions, Non_Annex_I_emissions)
```

## Export as a CSV file

```{r export as a csv file}
write.csv(UNFCCC_Emissions, "UNFCCC-Emissions-clean.csv")
```

## Export as an R script for future use

Only run this chunk manually once within the .Rmd file. It produces an error when knitting it as a whole because of chunk label duplicates. As of `r format(Sys.time(), '%B %d, %Y')`, there hasn't been a viable solution to run the code below when as part of the knitting process.


```{r export as an R script, eval=FALSE, message=FALSE, warning=FALSE}
knitr::purl("UNFCCC-Emissions.Rmd", "UNFCCC-Emissions-clean.R")
```
